---
title: "Class 2-8-15"
author: "Alan T. Arnholt"
date: "February 8, 2016"
output: 
  html_document: 
    highlight: textmate
    theme: readable
---

```{r, echo = FALSE, results= 'hide', message = FALSE, warning = FALSE}
set.seed(123)
library(knitr)
knitr::opts_chunk$set(comment = NA, fig.align = 'center', fig.height = 5, fig.width = 5, prompt = FALSE, highlight = TRUE, tidy = FALSE, warning = FALSE, message = FALSE, tidy.opts=list(blank = TRUE, width.cutoff= 75, cache = FALSE))
```

# Chapter 5

## Changing Values in Place

```{r}
vec <- c(0, 0, 0, 0, 0, 0)
vec
vec[1]
vec[1] <- 1000
vec
vec[c(1, 3, 5)] <- c(1, 1, 1)
vec
vec[4:6] <- vec[4:6] + 1
vec
vec[7] <- 0
vec
```

Start by creating a deck of cards:

```{r}
FACE <- c("king", "queen", "jack", "ten", "nine", "eight", "seven", "six", "five", "four", "three", "two", "ace")
SUIT <- c("spades", "clubs", "diamonds", "hearts")
VALUE <- 13:1
deck <- data.frame(face = rep(FACE, 4), suit = rep(SUIT, each = 13), value = rep(VALUE, times = 4))
library(DT)
datatable(deck)
```

### Add a Variable

```{r}
deck2 <- deck
deck2$new <- 1:52
datatable(deck2)
deck2$new <- NULL
head(deck2)
```

## Change Particular Values

```{r}
deck2[c(13, 26, 39, 52), ]
deck2[c(13, 26, 39, 52), "value"]
deck2[c(13, 26, 39, 52), 3]
deck2$value[c(13, 26, 39, 52)]
deck2$value[c(13, 26, 39, 52)] <- c(14, 14, 14, 14)
deck2$value[c(13, 26, 39, 52)]
# OR
deck2$value[c(13, 26, 39, 52)] <- 14
deck2$value[c(13, 26, 39, 52)]
```

## Recall `shuffle()`

This function still need some work...

```{r}
shuffle <- function(cards) {
 random <- sample(1:52, size = 52)
 cards[random, ]
}
deck3 <- shuffle(deck)
datatable(deck3)
```

## Logical Operators

```{r}
1 > 2
1 > c(0, 1, 2)
c(1, 2, 3) == c(3, 2, 1)
1 %in% c(3, 4, 5)
c(1, 2) %in% c(3, 4, 5)
c(1, 2, 3) %in% c(3, 4, 5)
c(1, 2, 3, 4) %in% c(3, 4, 5)
```

## Exercise

Extract the face column of `deck2` and test whether each value is equal to ace.  As a challenge, use `R` to quickly count how many cards are equal to `ace`.

```{r}
sum(deck2$face == "ace")
deck3$face == "ace"
```

Change the value of `ace` to 14

```{r}
deck3$value[deck3$face == "ace"] <- 14
datatable(deck3)
deck4 <- deck
```

## Exercise

Assign a value of 1 to every card in `deck4` that has a suit of hearts.

```{r}
deck4$value[deck4$suit == "hearts"] <- 1
deck4$value[deck4$suit == "hearts"]
datatable(deck4)
```

In hearts, the queen of spades has the most unusual value of all: she’s worth 13 points.

```{r}
deck4$value[deck4$suit == "spades" & deck4$face == "queen"] <- 13
deck4$value[deck4$suit == "spades" & deck4$face == "queen"]
datatable(deck4)
```

## Exercise

```{r}
w <- c(-1, 0, 1)
x <- c(5, 15)
y <- "February"
z <- c("Monday", "Tuesday", "Friday")
w > 0 
10 < x & x < 20
y == "February"
all(z %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))
deck5 <- deck
```

Change the value of all face cards to 10

```{r}
deck5$value[deck5$face %in% c("king", "queen", "jack")] <- 10
datatable(deck5)
```

## Missing Information

```{r}
1 + NA
NA == 1
c(NA, 1:50)
mean(c(NA, 1:50))
mean(c(NA, 1:50), na.rm = TRUE)
NA == NA
c(1, 2, 3, NA) == NA
is.na(NA)
vec <- c(1, 2, 3, NA)
is.na(vec)
deck5$value[deck5$face == "ace"] <- NA
head(deck5, 13)
```

```{r, echo = FALSE, results= 'hide', message = FALSE, warning = FALSE}
set.seed(123)
library(knitr)
knitr::opts_chunk$set(comment = NA, fig.align = 'center', fig.height = 5, fig.width = 5, prompt = FALSE, highlight = TRUE, tidy = FALSE, warning = FALSE, message = FALSE, tidy.opts=list(blank = TRUE, width.cutoff= 75, cache = FALSE))
```


# Environments

```{r}
library(pryr)
library(knitr)
parenvs(all = TRUE)
as.environment("package:knitr")
globalenv()
baseenv()
emptyenv()
```

Next, you can look up an environment's parent with `parent.env`:

```{r}
parent.env(globalenv())
```

Notice that the empty environment is the only R environment without a parent:

```{r, error = TRUE}
parent.env(emptyenv())
```

You can view the objects saved in an environment with `ls` or `ls.str`. `ls` will return just
the object names, but `ls.str` will display a little about each object’s structure:

```{r}
ls(emptyenv())
ls(globalenv())
```

You can use `R`’s `$` syntax to access an object in a specific environment. For example, you can access deck from the global environment:

```{r}
head(globalenv()$deck)
```

And you can use the assign function to save an object into a particular environment.  First give assign the name of the new object (as a character string). Then give assign the value of the new object, and finally the environment to save the object in:

```{r}
assign("NewObject", "Value of new object", envir = globalenv())
globalenv()$NewObject
```

You can use `environment` to see the current active environment:

```{r}
environment()
```

## Evaluation

`R` creates a new environment each time it evaluates a function. `R` will use the new environment as the active environment while it runs the function, and then `R` will return to the environment that you called the function from, bringing the function’s result with it. Let’s call these new environments runtime environments because `R` creates them at runtime to evaluate functions.  We’ll use the following function to explore `R`’s runtime environments. We want to know what the environments look like: what are their parent environments, and what objects do they contain? `show_env` is designed to tell us:

```{r}
show_env <- function(){
  list(ran.in = environment(),
       parent = parent.env(environment()),
       objects = ls.str(environment()))
}
```

`show_env` is itself a function, so when we call `show_env()`, `R` will create a runtime environment to evaluate the function in. The results of `show_env` will tell us the name of the runtime environment, its parent, and which objects the runtime environment contains:


```{r}
a <- show_env()
a
b <- a$ran.in
b
```

The results reveal that `R` created a new environment to run
`show_env()` in. The environment had no objects in it, and its parent was the global
environment. 

Let's run `show_env` again:

```{r}
show_env()
```

You can look up a function's origin environment by running `environment` on the function:

```{r}
environment(PASWR2::z.test)
environment(PASWR::z.test)
environment(BSDA::z.test)
environment(show_env)
```



## The Functions

```{r}
setup <- function(deck) {
  DECK <- deck
  DEAL <- function() {
    card <- deck[1, ]
    assign("deck", deck[-1, ], envir = parent.env(environment()))
    card
  }
  SHUFFLE <- function(){
    random <- sample(1:52, size = 52)
    assign("deck", DECK[random, ], envir = parent.env(environment()))
  }
  list(deal = DEAL, shuffle = SHUFFLE)
}
cards <- setup(deck)
deal <- cards$deal
shuffle <- cards$shuffle
```

## Where Is All of This?

```{r}
setup
environment(setup)
deal
environment(deal)
shuffle
environment(shuffle)
```

## Playing Cards Now

```{r}
shuffle()
deal()
deal()
deal()
deal()
deal()
```