---
title: "Chapters 1 and 2 Notes"
author: "Alan Arnholt"
date: "2/15/2016"
output: 
  html_document: 
    highlight: textmate
    theme: readable
---

```{r, echo = FALSE, results= 'hide', message = FALSE, warning = FALSE}
set.seed(123)
library(knitr)
knitr::opts_chunk$set(comment = NA, fig.align = 'center', fig.height = 5, fig.width = 5, prompt = FALSE, highlight = TRUE, tidy = FALSE, warning = FALSE, message = FALSE, tidy.opts=list(blank = TRUE, width.cutoff= 75, cache = FALSE))
```


The outer product $\boldsymbol{u} \otimes \boldsymbol{v}^T$ is equivalent to a matrix multiplication $\boldsymbol{u}\boldsymbol{v}^T$, provided $\boldsymbol{u}$ is represented as a $m \times 1$ column vector and $\boldsymbol{v}$ as a $n \times$ column vector.  See [Wikipedia](https://en.wikipedia.org/wiki/Outer_product) for more details.

```{r}
u <- 1:4
v <- 1:3
uOUTERv <- u %o% v
uOUTERv
```

The inner product $\langle \boldsymbol{u}, \boldsymbol{v}\rangle = \boldsymbol{u}^T\boldsymbol{v}$ provided $m=n$ yields a scalar (or $1 \times 1$ matrix).

```{r}
u <- 1:6
v <- 1:6
u %*% v
sum(u^2)
```

## Functions

When working with nested functions, operations all start with the innermost operation and work to the outermost.

```{r}
set.seed(3)
round(mean(sample(1:10)))
```

## Sample with Replacement

```{r}
set.seed(4)
sample(1:6, size = 2, replace = FALSE) # replace = FALSE is the default
sample(1:6, size = 2, replace = TRUE)
```

## Writing Your Own Functions

```{r}
roll <- function(){
  die <- 1:6
  dice <- sample(die, size = 2, replace = TRUE)
  sum(dice)
}
roll()
roll()
roll()
```

```{r}
roll2 <- function(bones){
  dice <- sample(bones, size = 2, replace = TRUE)
  sum(dice)
}
roll2(bones = 1:4)
roll2(bones = 1:6)
roll2(bones = 1:10)
```

## Supplying Default Arguments

```{r}
roll2 <- function(bones = 1:6){
  dice <- sample(bones, size = 2, replace = TRUE)
  sum(dice)
}
roll2()
```

## Chapter 2

### `ggplot2`

#### Scaterplots

```{r}
x <- seq(from = -1, to = 1, by  = 0.2)
x
y <- x^3
DF <- data.frame(x = x, y = y)
library(ggplot2)
ggplot(data = DF, aes(x = x, y = y)) + 
  geom_point() +
  theme_bw()
```

#### Histograms

```{r}
x <- c(1, 2, 2, 2, 3, 3)
DF2 <- data.frame(x = x)
ggplot(data = DF2, aes(x = x)) +
  geom_histogram(binwidth = 1, fill = "purple", color = "black") + 
  theme_bw()
```

### `ggvis`

#### Scatterplots

```{r, message = FALSE, warning = FALSE}
library(ggvis)
DF %>% 
  ggvis(x = ~x, y = ~y) %>%
  layer_points()
```

#### Histograms

```{r}
DF2 %>% 
  ggvis(x = ~x, fill := "green") %>% 
  layer_histograms(width = 1)
```

## Exercise

```{r}
x3 <- c(0, 1, 1, 2, 2, 2, 3, 3, 4)
DF3 <- data.frame(x = x3)
# Use ggplot2 first
ggplot(data = DF3, aes(x = x)) + 
  geom_histogram(binwidth = 1, fill = "red", color = "black") + 
  theme_bw()
# Use ggvis next
DF3 %>% 
  ggvis(x = ~x, fill := "lightblue") %>% 
  layer_histograms(width = 1)
```

### Using `replicate()`

```{r}
replicate(10, 3 + 3.24)
replicate(10, roll())
rolls <- replicate(10000, roll())
STR <- system.time(replicate(10000, roll()))
STR
DF4 <- data.frame(rolls = rolls)
# Use ggplot2 first
ggplot(data = DF4, aes(x = rolls)) +
  geom_histogram(binwidth = 1, fill = "wheat", color = "black") +
  theme_bw()
# Use ggvis next
DF4 %>% 
  ggvis(x = ~rolls, fill := "salmon") %>% 
  layer_histograms(width = 1)
```

### Another Approach to `replicate()`

```{r}
N <- 10000
rolls <- numeric(N)
for(i in 1:N){
  rolls[i] <- roll()
}
# Make this into a function
FL <- function(N = 10000){
  rolls <- numeric(N)
  for(i in 1:N){
  rolls[i] <- roll()
}
}
STL <- system.time(FL())
STL
DF5 <- data.frame(rolls = rolls)
# Use ggplot2 first
ggplot(data = DF5, aes(x = rolls)) +
  geom_histogram(binwidth = 1, fill = "purple", color = "black") +
  theme_bw()
# Use ggvis next
DF5 %>% 
  ggvis(x = ~rolls, fill := "indianred") %>% 
  layer_histograms(width = 1)
```

## Exercise

```{r}
roll <- function(){
  die <- 1:6
  dice <- sample(die, size = 2, replace = TRUE, 
                 prob = c(1/8, 1/8, 1/8, 1/8, 1/8, 3/8))
  sum(dice)
}
rolls <- replicate(10000, roll())
DF6 <- data.frame(rolls = rolls)
# Use ggplot2 first
ggplot(data = DF6, aes(x = rolls)) +
  geom_histogram(binwidth = 1, fill = "steelblue", color = "black") +
  theme_bw()
# Use ggvis next
DF6 %>% 
  ggvis(x = ~rolls, fill := "magenta") %>% 
  layer_histograms(width = 1)
```

## Uber Die

```{r}
roll <- function(dietype = 1:6, probvalues = rep(1/length(dietype), length(dietype))){
  if(sum(probvalues) != 1)
    stop("'probvalues' must add to one")
  dice <- sample(dietype, size = 2, replace = TRUE, prob = probvalues)
  sum(dice)
}
roll(dietype = c(1,3,5,7))
rolls <- replicate(10000, roll(dietype = 1:20))
DF7 <- data.frame(rolls = rolls)
# Use ggplot2 first
ggplot(data = DF7, aes(x = rolls)) +
  geom_histogram(binwidth = 1, fill = "lightblue", color = "black") +
  theme_bw()
# Use ggvis next
DF7 %>% 
  ggvis(x = ~rolls, fill := "orange") %>% 
  layer_histograms(width = 1)
```
