---
title: "Chapters 3 and 4"
author: "Alan Arnholt"
date: 'Last Updated: `r Sys.Date()`'
output: html_document
---

```{r, label = "SETUP", echo = FALSE, results= 'hide', message = FALSE, warning = FALSE}
knitr::opts_chunk$set(comment = NA, fig.align = 'center', fig.height = 5, fig.width = 5,  warning = FALSE, message = FALSE, tidy.opts=list(blank = TRUE, width.cutoff = 75))
```

# Chapter 3 Notes

## Exercise

Can you spot the difference between a character string and a number?  Here's a test: Which of these are character strings and which are numbers? 1, "1", "one".

```{r}
x <- 1
y <- "1"
z <- "one"
typeof(1)
typeof(x)
typeof("1")
typeof(y)
typeof("one")
typeof(z)
```

## Exercise

Create an atomic vector that stores just the face names of the cards in a royal flush (ace through ten all one suit), for example, the ace of spades, king of spades, queen of spades, jack of spades, and ten of spades.  The face name of the ace of spades would be "ace", and "spades" is the suit.  Which type of vector will you use to save the names?

```{r}
FaceNames <- c("ace", "king", "queen", "jack", "ten")
FaceNames
typeof(FaceNames)
```

## Exercise

Create the following matrix, which stores the name and suit of every card in a royal flush.

```{r echo = FALSE}
MAT <- matrix(data = c(FaceNames, rep("spaces", 5)), ncol = 2)
MAT
```

```{r}
hand1 <- c("ace", "king", "queen", "jack", "ten", rep("spades", 5))
matrix(hand1, ncol = 2)
matrix(hand1, nrow = 5)
dim(hand1) <- c(5, 2)
hand1
```

**Note:** R matrices are column major.

## Dates and Times

```{r}
now <- Sys.time()
now
typeof(now)
class(now)
SEC <- unclass(now)
SEC
```

The number stored in `SEC` represents the number of seconds that have passed between the time and 12:00 AM January 1st 1970 (in the Universal Time Coordinated (UTC) zone).

## Factors

```{r}
gender <- factor(c("male", "female", "female", "male"))
typeof(gender)
attributes(gender)
unclass(gender)
```

## Exercise


Many card games assign a numerical value to each card.  For example, in blackjack, each face card is worth 10 points, each number card is worth between 2 and 10 points, and each ace is worth 1 or 11 points, depending on the final score.  Make a virtual card by combining "ace", "heart", and 1 into a vector.  What type of atomic vector will result?  **Character**  Check if you are right.

```{r}
card <- c("ace", "heart", 1)
typeof(card)
```

## Exercise

Use a list to store a single playing card, like the ace of hearts, which has a point value of one.  The list should save the face of the card, the suit, and the point value in separate elements.

```{r}
card <-  list(face = "ace", suit = "hearts", value = 1)
card
```

## Data Frames

```{r}
df <- data.frame(face = c("ace", "two", "six"), suit = c("clubs", "clubs", "clubs"), value = c(1, 2, 3))
df
typeof(df)
class(df)
str(df)
df2 <- data.frame(face = c("ace", "two", "six"), suit = c("clubs", "clubs", "clubs"), value = c(1, 2, 3), stringsAsFactors = FALSE)
df2
typeof(df2)
class(df2)
str(df2)
```

**Creating a deck of cards with less typing**

```{r}
Face <- c("king", "queen","jack", "ten","nine","eight","seven","six","five","four","three","two","ace")
Suit <- c("spades","clubs", "diamonds", "hearts")
Value = 13:1
deck <- data.frame(face = rep(Face, 4), suit = rep(Suit, each = 13), value = rep(Value, 4), stringsAsFactors = FALSE)
library(DT)
datatable(deck)
```

## Reading a file from a secure web site using `readr::read_csv()` and `repmis::source_data()`

**Note:** `read.csv()` will not read from `https` (Hypertext Transfer Protocol Secure) web sites.

```{r message = TRUE}
site <- "https://gist.githubusercontent.com/garrettgman/9629323/raw/ee5dfc039fd581cb467cc69c226ea2524913c3d8/deck.csv"
deck2 <- readr::read_csv(site)
head(deck2)
deck1 <- repmis::source_data(url = site, sep = ",", header = TRUE)
datatable(deck1)
```

## Saving Data

```{r}
write.csv(deck1, file = "cards.csv", row.names = FALSE)
```

## Downloading any file secure or not

We download the `deck.csv` file from the supplied GitHub url stored in the character string `site` and store the downloaded file as `DFcards.csv` in the same directory as the current document.

```{r}
download.file(url = site, destfile = "./DFcards.csv", method = "curl")
list.files()
```

# Chapter 4 Notes

## Selecting Values

```{r}
head(deck)
deck[1, 1]
deck[1, 1:3]
deck[1:2, 1]  # returns a single column
deck[1:2, 1, drop = FALSE] # returns a data frame
deck[-(2:52), 1:3] # everything except rows 2-52 for cols 1-3
```

### Blank Spaces

You can use a blank space to tell `R` to extract _every_ value in a dimension.

```{r}
deck[1, ]  # same as deck[1, 1:3]
```

### Logical Values

```{r}
set.seed(4)
sims <- 10000
xbar <- numeric(sims)
for(i in 1:sims){
  xbar[i] <- mean(runif(50, 0, 10))
}
mean(xbar)
sd(xbar)
library(ggplot2)
DF <- data.frame(xbar = xbar)
ggplot(data = DF, aes(x = xbar)) + 
  geom_density(fill = "lightblue") + 
  stat_function(fun = dnorm, args = list(mean = 5, sd = ((10)/sqrt(12))/sqrt(50)), color = "red") + 
  theme_bw()
```

What percent of the values in `xbar` are between $5 - (10/\sqrt{12})/(\sqrt{50}) = `r (5 - ((10)/sqrt(12))/sqrt(50))`$ and $5 + (10/\sqrt{12})/(\sqrt{50}) = `r (5 + ((10)/sqrt(12))/sqrt(50))`$?

```{r}
LOG <- xbar >= (5 - ((10)/sqrt(12))/sqrt(50)) & xbar <= (5 + ((10)/sqrt(12))/sqrt(50))
head(LOG, n = 10)
head(which(LOG == TRUE), n = 10)
length(xbar[LOG])
mean(LOG)
# Compare to Normal
pnorm(1) - pnorm(-1)
```

## Exercise

Use the preceding ideas to write a shuffle function.  `shuffle()` should take a data frame and return a shuffled copy of the data frame.

```{r}
shuffle <- function(cards){
  index <- sample(dim(cards)[1], size = dim(cards)[1], replace = FALSE)
  cards[index, ]
}
deck2 <- shuffle(cards = deck)
deck2[1:5, ]
```

## Dollar Signs and Double Brackets

```{r}
deck$value
mean(deck$value)
LST <- list(numbers = c(1, 2, 3, 4, 5), logical = c(TRUE, FALSE, TRUE), strings = c("dog", "cat", "horse", "car"))
LST
```

Subsetting the first element:

```{r}
LST[1]             # a list
LST[[1]]           # values inside element
LST$numbers        # values inside element
LST[['numbers']]   # values inside element
```

If you subset a list with single-bracket notation, `R` will return a smaller list.  If you subset a list with double-bracket notation, `R` will return the values inside the element subsetted of the list.
