---
title: "Creating Data Frames from Contingency Tables"
author: "Alan Arnholt"
date: 'Last Updated: `r Sys.Date()`'
output: 
  html_document: 
    highlight: textmate
    theme: readable
---

```{r, label = "SETUP", echo = FALSE, results= 'hide', message = FALSE, warning = FALSE}
knitr::opts_chunk$set(comment = NA, fig.align = 'center', fig.height = 5, fig.width = 5,  warning = FALSE, message = FALSE, tidy.opts=list(blank = TRUE, width.cutoff = 75))
```

Journal articles and books often present information in contingency tables.  This document briefly considers how to store the information from a contingency table as a data frame.  Consider the contingency table `mat` stored as a matrix.

```{r}
mat <- matrix(data = c(56, 35, 61, 43, 54, 61, 21, 42, 8, 19), nrow = 2)
dimnames(mat) <- list(gender = c("girls", "boys"), ability = c("hopeless", "belowavg", "average", "aboveavg", "superior"))
mat
class(mat)
```

One approach to create a data frame from a contingency table is to use the `expand.dft` function from the `vcdExtra` package.

1. If the data is stored in a matrix, convert it to class `table`

```{r}
matT <- as.table(mat)
matT
class(matT)
```

2. Convert the table object to a data frame using `as.data.frame`.

```{r}
matDF <- as.data.frame(matT)
matDF
class(matDF)
```

3. Convert the frequency form data frame to a data frame using the `expand.dft` function.

```{r}
DF <- vcdExtra::expand.dft(matDF)
head(DF)
class(DF)
```

Consider creating a contingency table from the data frame `DF`.

```{r}
CT <- xtabs(~gender + ability, data = DF)
CT
chisq.test(CT)
```

## Randomization Test

```{r}
set.seed(2)
N <- 10^4 - 1 # Change this for slower computers
result <- numeric(N)
for (i in 1:N) {
T2 <- xtabs(~ sample(gender) + ability, data = DF)
result[i] <- chisq.test(T2)$statistic
}
obs <- chisq.test(xtabs(~ gender + ability, data = DF))$statistic
obs
pvalue <- (sum(result >= obs) + 1)/(N + 1)
pvalue
```



