---
title: "Object Oriented Programming"
author: "Alan Arnholt"
date: 'Last updated: `r format(Sys.time(), "%b %d, %Y")`'
output: html_document
---

```{r, label = "SETUP", echo = FALSE, results= 'hide', message = FALSE, warning = FALSE}
library(knitr)
knitr::opts_chunk$set(comment = NA, fig.align = 'center', fig.height = 5, fig.width = 5, prompt = FALSE, highlight = TRUE, tidy = FALSE, warning = FALSE, message = FALSE, tidy.opts=list(blank = TRUE, width.cutoff = 80))
```

## R Themes^[The Art of R Programming]

* Everything you touch in R - ranging from numbers to character strings to matrices - is an object.

* R promotes *encapsulation*, which is packaging separate but related data items into one class instance.  Encapsulation helps you keep track of related variables, enhancing clarity.

* R classes are *polymorphic*, which means that the same function call leads to different operations for objects of different classes.  For instance, a call to `print()` on an object of a certain class triggers a call to a print function tailored to that class.  Polymorphism promotes usability.



## S3 Classes

An S3 class consists of a list, with a class name attribute and *dispatch* capability added.

* `class` is an attribute of an object that dictates what message the object can receive and return.
* `method` is a function that is designed for a specific class.
* `dispatch` is selection of the class specific method.

## S3 Generic Functions

When a generic function is called, R will dispatch the call to the proper class method, meaning that it will reroute the call to a function defined for the object's class.

1.  A generic function (`plot`, `print`, `summary`, etc.) checks the class of the object.
2.  A search is done to see if there is an appropriate method for that class.
3.  If there exists a method for that class, then that method is called on the object and we are finished.
4.  If a method for a class does not exist, a search is done to see if there is a `default` method for the generic function.  If `default` exists, then it is called.
5.  If no default exists an error appears.



```{r}
plot
```

Note that the generic function `plot` has `UseMethod("plot")` in its function call.  To see all implementations of a generic function type `methods(GenericFunction)`.

```{r}
methods(plot)
```

One way to view the code for a hidden method (one that ends with `*`) is to use the `getAnywhere` function.

```{r}
getAnywhere(plot.density)
```

## Writing S3 Classes

A class instance is created by forming a list, with the components of the list being the member variables of the class.  The `class` attribute is set by hand using the `class` function, and then various implementations of generic functions are defined.  Consider the following slightly modified code taken from *S Programming* for a 2 sample equal variance t-test.

Recall that the formula for the 2 sample equal variance t-test is:

$$t_{obs} = \frac{(\bar{x} - \bar{y}) - (\mu_x - \mu_y)}{\sqrt{1/n_x + 1/n_y}\cdot \sqrt{\left((n_x-1)\cdot s_x^2 + (n_y-1)\cdot s_y^2)\right)/(n_x + n_y - 2)}} \sim t_{n_x + n_y -2}$$

```{r}
ttest <- function(x, y, alpha = 1/20, alternative = "two.sided", ...){
  if(is.null(class(x))){
    class(x) <- data.class(x)
  }
  UseMethod("ttest")
}
```

The `ttest.default` is written next.  Note that a list (`results`) is constructed and the given the class `my.t.test` with the code `class(results) <- "my.t.test"`.

```{r}
ttest.default <- function(x, y, alpha = 1/20, alternative = "two.sided", ...){
  choices <- c("two.sided", "greater", "less")
  alt <- pmatch(alternative, choices)
  alternative <- choices[alt]
  # add some checks here
  yok <- !is.na(y)
  xok <- !is.na(x)
  x <- x[xok]
  y <- y[yok]
  n1 <- length(x)
  n2 <- length(y)
  ndf <- n1 + n2 - 2
  s2 <- ((n1 - 1)*var(x) + (n2 - 1)*var(y))/ndf
  tstat <- (mean(x) - mean(y))/sqrt(s2*(1/n1 + 1/n2))
  # tails
  if(alternative == "two.sided"){
    tail.area <- 2 * (1 - pt(abs(tstat), ndf))
  } else if(alternative == "less") {
    tail.area <- pt(tstat, ndf)
  } else {
    tail.area <- 1 - pt(tstat, ndf)
  }
  results <- list("t-stat" = tstat, d.f. = ndf, y1 = x, y2 = y,
                  alternative = alternative, tail.area = tail.area,
                  reject = tail.area < alpha, alpha = alpha)
  class(results) <- "my.t.test"
  results
}
```

Objects of class `my.t.test` will be printed using the generic `print` after defining `print.my.t.test`.

```{r}
print.my.t.test <- function(x, ...){
  cat("\n")
  cat("My Monday Night Modified 2 Sample Equal Variance t-test", "\n")
  cat("\n")
  cat(paste("Alternative hypothesis:", x[5]), "\n")
  cat(paste("Reject the null hypothesis:", x[7]), "\n")
  cat("\n")
  print(round(unlist(x[c(1, 2, 6)]), 4))
  invisible(x)
}
```

Objects of class `my.t.test` will be plotted using the generic `plot` after defining `plot.my.t.test`.
```{r}
plot.my.t.test <- function(x, ...){
  boxplot(x[c("y1", "y2")], col = "pink",
          names = c("Sample 1", "Sample 2"), ...)
}
```

Additional methods 

```{r}
ttest.matrix <- function(x, ...){
  ttest(x[, 1], x[, 2])
}

ttest.list <- function(x, ...){
  ttest(x = x[[1]], y = x[[2]])
}
```

## Does This Work?

```{r}
set.seed(123)
TT <- ttest(rnorm(100, 0, 1), rnorm(100, 1, 1))
TT 
print(TT)
plot(TT)
Mx1x2 <- matrix(rnorm(60, 0, 1), nrow = 30)
class(Mx1x2)
TM <- ttest(Mx1x2)
TM
plot(TM)
mylist <- list(xl = rnorm(30), yl = rnorm(30))
class(mylist)
ttest(mylist)
plot(ttest(mylist))
```

## How To Package?

What does the `ttest.R` file with `roxygen` look like?

```
#' Two-Sample Equal Variance t-Test
#'
#' @param x a non-empty numeric vector.
#' @param y a non-empty numeric vector.
#' @param alpha value to test hypothesis.
#' @param alternative hypothesis (can take on one of three values).
#' @param ... further arguments to be passed to or from methods.
#'
#' @return the answers!
#' @export
#'
#' @examples
#' T1 <- ttest(x = rnorm(300.5, 1), y = rnorm(30, 1, 1), alternative = "less")
#' T1
#' print(T1)
#' plot(T1)
#' Mx1x2 <- matrix(rnorm(60, 0, 1), nrow = 30)
#' class(Mx1x2)
#' TM <- ttest(x = Mx1x2)
#' TM
#' plot(TM)
#' ttest(x = Mx1x2) # Test that a matrix will work
#' mylist <- list(xl = rnorm(30), yl = rnorm(30))
#' class(mylist)
#' ttest(x = mylist, alternative = "greater")
#' plot(ttest(mylist))
#'
ttest <- function(x, y, alpha = 1/20, alternative = "two.sided", ...){
  if(is.null(class(x))){
    class(x) <- data.class(x)
  }
  UseMethod("ttest")
}



#' @export
ttest.default <- function(x, y, alpha = 1/20, alternative = "two.sided", ...){
  choices <- c("two.sided", "greater", "less")
  alt <- pmatch(alternative, choices)
  alternative <- choices[alt]
  # add some checks here
  yok <- !is.na(y)
  xok <- !is.na(x)
  x <- x[xok]
  y <- y[yok]
  n1 <- length(x)
  n2 <- length(y)
  ndf <- n1 + n2 - 2
  s2 <- ((n1 - 1)*var(x) + (n2 - 1)*var(y))/ndf
  tstat <- (mean(x) - mean(y))/sqrt(s2*(1/n1 + 1/n2))
  # tails
  if(alternative == "two.sided"){
    tail.area <- 2 * (1 - pt(abs(tstat), ndf))
  } else if(alternative == "less") {
    tail.area <- pt(tstat, ndf)
  } else {
    tail.area <- 1 - pt(tstat, ndf)
  }
  results <- list("t-stat" = tstat, d.f. = ndf, y1 = x, y2 = y,
                  alternative = alternative, tail.area = tail.area,
                  reject = tail.area < alpha, alpha = alpha)
  class(results) <- "my.t.test"
  results
}

#' @export
print.my.t.test <- function(x, ...){
  cat("\n")
  cat("My Monday Night Modified 2 Sample Equal Variance t-test", "\n")
  cat("\n")
  cat(paste("Alternative hypothesis:", x[5]), "\n")
  cat(paste("Reject the null hypothesis:", x[7]), "\n")
  cat("\n")
  print(round(unlist(x[c(1, 2, 6)]), 4))
  invisible(x)
}

#' @export
plot.my.t.test <- function(x, ...){
  boxplot(x[c("y1", "y2")], col = "pink",
          names = c("Sample 1", "Sample 2"), ...)
}

#' @export
ttest.matrix <- function(x, ...){
  ttest(x[, 1], x[, 2])
}

#' @export
ttest.list <- function(x, ...){
  ttest(x = x[[1]], y = x[[2]])
}
```

Note: I documented the function `ttest` and exported `ttest` to the `namespace` using the tag `export`.  The various methods for `ttest` are exported using the `@export` tag but are not documented (my choice).

```{r}
methods(ttest)
```

```{r}
methods(print)[111:120]
```

```{r}
methods(plot)
```

```{r}
getAnywhere(print.my.t.test)
```